---
title: "Statistical Analysis"
output: pdf_document
date: "2024-10-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
video_data <- read.table("videodata.txt", header = TRUE, sep="")
video_multiple <- read.table("videoMultiple.txt", header = TRUE, sep="")
```

## Contribution

# 1. Introduction

This report analyzes a survey conducted among 3000 - 4000 students enrolled in statistics courses at UC Berkeley during the Fall 1994 semester. The survey aimed to investigate students' video game habits, preferences, and attitudes towards gaming. A total of 314 students from Statistics 2, Section 1 were eligible for the survey based on their participation in the second exam of the semester, with 95 students randomly selected to participate. Out of the selected group, 91 students submitted complete responses, with incomplete responses encoded as 99.

The survey covered questions including the number of hours spent playing video games, where and how often students play, their preferences for different types of games, etc. The second part of the survey covered students' attitudes toward video games, whether they liked playing games or not and why.

The primary goal of this analysis is to investigate students' video game behaviors and preferences, as well as potential correlations between these behaviors and demographic factors such as gender, work, and computer ownership. Specifically, we aim to investigate the proportion of students playing video games during the week prior to the survey, the relation between reported time spend playing video games, the frequency of playing, and timing of the survey, the main reasons students like or dislike playing video games, and any noticeable differences in video game preferences and habits based on demographics. We found that **insert results of analysis here**

The report is structured as follows: Section 1 addresses each of the research questions outlined above, Section 2 analyzes the survey data, including numerical and graphical comparisons of video game habits and preferences. Section 3 expands on the analysis to explore potential differences across subgroups. Section 4 summarizes the findings and discusses their implications, and Section 5 provides additional information and appendices related to the analysis.

# 2. Analysis

## 2.1. Estimate Proportions of Students Who Played a Video Game

### 2.1.1. Methods

The first step of our analysis was to estimate the proportion of students who played a video in the week prior to the survey. We found both a point estimate and an interval estimate for this proportion.

The point estimate represents a single estimated value of the proportion of students who played video games based on the sample. The sample proportion \( \hat{p} \) is calculated as the ratio of students who played video games (i.e., reported playing for more than 0 hours) to the total number of students surveyed.
   
   The formula for the sample proportion is given by:
   \[
   \hat{X} = \frac{k}{n}
   \]
   where \( k \) is the number of students who played video games, and \( n \) is the total number of students surveyed.

The interval estimate provides a range that is likely to contain the true proportion of students who played video games. This is calculated using a **confidence interval with a 95% confidence interval.

   The confidence interval is calculated as:
   \[
   (\hat{X} - 2 \frac{\sigma}{\sqrt{n}}, \hat{X} + 2 \frac{\sigma}{\sqrt{n}})
   \]
   where \(\frac{\sigma}{\sqrt{n}} \) is the standard error.

### 2.1.2. Analysis

To find the point estimate, we found the sample proportion by counting the number of students whose playing time a week prior to the survey was greater than zero and the total number of responses to the survey. The proportion of students who played video games in the week prior to the survey is approximately 37.36%. While the point estimate gives a specific value, it does not account for the inherent variability that comes with sampling. To take this uncertainty into account, we calculated a 95% confidence interval. The 95% confidence interval for the proportion of students who played video games is between 27.22% and 47.51%.

The point estimate provides a single value representing the proportion of students who played video games, while the confidence interval provides a range that accounts for uncertainty due to sampling. The interval is based on the assumption that our sample is representative of the entire population of students who took the survey. The confidence interval ensures that if the survey were repeated multiple times, 95% of the intervals calculated in this manner would contain the true population proportion.

### 2.1.3. Conclusion

A point estimator is a single value $\bar{x}$, useful for summarizing the central tendency. In this case, it estimates the proportion of students that played video games a week prior to the survey given a sample. An interval estimate is a range of two values, where one can estimate that the population proportion is between those two values with a fixed confidence (probability), reflecting the variability in the sample data. In our case, the interval estimate is very wide, reducing its usefulness in estimating the population proportion, since the true proportion could plausibly fall anywhere between 27% and 47%.

## 2.2. Comparison of Time Spent and Reported Frequency of Playing Video Games

### 2.2.1. Methods

Next, we wanted to look at the if the amount of time spent playing games in the week prior to the survey compares for different reported frequencies. In the survey, students were asked to report how often they play video games on a scale from 1-4, where 1 represents playing daily, 2 represents playing weekly, 3 represents playing monthly, and 4 represents playing semesterly. To compare the amount of time spent playing games between different reported frequencies, we graphed the distributions of time spent playing games and calculated the proportion of students who reported playing video games in each frequency category and the average time they spent playing. This comparison helped us assess whether students' reported gaming habits align with the actual hours they spent playing video games. We considered the potential impact of an exam held the week before the survey, which may have influenced the amount of time students spent playing video games that week.

### 2.2.2. Analysis

```{r, echo=FALSE, fig.width=3.5, fig.height=2.5}
daily_gamers <- video_data[video_data$freq == 1,]
weekly_gamers <- video_data[video_data$freq == 2,]
monthly_gamers <- video_data[video_data$freq == 3,]
semesterly_gamers <- video_data[video_data$freq == 4,]

hist(daily_gamers$time, 
     main = "Daily Gamers", 
     xlab = "Hours Played",
     breaks=20)
hist(weekly_gamers$time, 
     main = "Weekly Gamers", 
     xlab = "Hours Played",
     breaks=20)
hist(monthly_gamers$time, 
     main = "Monthly Gamers", 
     xlab = "Hours Played",
     breaks=20)
hist(semesterly_gamers$time, 
     main = "Semesterly Gamers", 
     xlab = "Hours Played",
     breaks=20)
```
Based on these graphs, it becomes apparent that daily and weekly gamers spent a lot more time playing games a week prior to this survey compared to monthly and semesterly gamers.  


```{r, echo=FALSE}
library(knitr)
library(kableExtra)

numerical_summaries <- data.frame(
  Sample_Proportions = c('77.78%', '85.71%', '11.11%', '4.35%')
)

rownames(numerical_summaries) <- c("Daily", "Weekly", "Monthly", "Semesterly")

kable(numerical_summaries, caption = "Sample Proportions Based On Reported Frequency") %>%
  kable_styling(latex_options = c("striped", "hold_position"), full_width = FALSE)
```

Similarly, the numerical summaries show that more daily and weekly gamers spent time playing games the week prior to the survey than monthly and semesterly gamers.

### 2.2.3. Conclusion

This survey was taken a week after a statistics exam, and the participants were students who participated in that exam. The fact that there was an exam in the week prior to the survey could possibly affect the previous estimates and this comparison. Students who reported playing daily or weekly generally followed through with significant gaming time during the week prior to the survey, seen by their higher point estimates and the shape of their time distributions. This suggests a strong alignment between reported gaming habits and actual behavior for these groups. On the otherhand, the majority of students who reported playing monthly or semesterly did not play video games during the week prior to the survey. Their self-reported frequency of playing (monthly or semesterly) may correctly reflect their long-term habits, however the timing of the survey (after the exam) may have influenced their decision not to play games during that specific week, which could explain the low percentages of gameplay in those groups. This could also explain why the overall proportion of students who played games in the past week was  lower than expected. Other variables in the survey like "Hate Math", "Grade Expected", or "Play if Busy" may be confounding variables that affect the results of our estimation. For example, those that respond "Yes" to the question "Hate Math" may dedicate less time to studying and more time to playing games. Similarly, those that already have high expectations for what grade they would earn on the upcoming exam might spend less time studying and more time playing games. Those that respond "Yes" to "Play if Busy" may still decide to spend time playing games despite being busy studying for the exam.

## 2.3. Estimate Means of Time Spent Playing

### 2.3.1. Methods

### 2.3.2. Analysis

### 2.3.3. Conclusion

# 3. Advanced Analysis

# 4. Discussion

# 5. Appendix
